# Identity By State (IBS) for 1000 Genomes BRCA1

```{r echo=FALSE, eval=FALSE}
# This codelab assumes that the current working directory is where the Rmd file
# resides
setwd("/YOUR/PATH/TO/codelabs/R/1000Genomes-BRCA1-analysis/data/plinkseqIBS/brca1")
```

The input IBS matrix is an N * N matrix, where N is the size of the
population. First, transform this matrix into a N^2 * 3 matrix, where
each row represents the IBS score for a pair of individuals.

```{r message=FALSE, comment=NA}
library(reshape2)
ibsData <- read.table("./1000genomes_phase1_brca1.ibs", header=TRUE,
                       row.names=1, stringsAsFactors=FALSE)
ibsMatrix <- data.matrix(ibsData)
ibsPairs <- melt(ibsMatrix)
colnames(ibsPairs) <- c("sample1", "sample2", "ibsScore")
write.table(ibsPairs, file="1000genomes_phase1_brca1_melted.ibs")
```

Extract the IBS matrix for a random sample of individuals.

```{r message=FALSE, comment=NA}
individuals <- unique(ibsPairs$sample1)
sampleSize <- 50
sample <- sample(individuals, sampleSize)
ibsPairs <- subset(ibsPairs, ibsPairs$sample1 %in% sample)
ibsPairs <- subset(ibsPairs, ibsPairs$sample2 %in% sample)
```

Exclude the IBS values for a genome and itself, because that's always 1.0 and
skews the heat map.

```{r message=FALSE, comment=NA}
ibsPairs <- subset(ibsPairs, ibsPairs$sample1 != ibsPairs$sample2)
```
Draw a heat map based on the IBS scores.

```{r ibs-heat-map, fig.align="center", fig.width=10, fig.height=10, message=FALSE, comment=NA}
require(ggplot2)
p <- ggplot(data=ibsPairs, aes(x=sample1, y=sample2)) +
     theme(axis.ticks=element_blank(), axis.text=element_blank()) +
     geom_tile(aes(fill=ibsScore), colour="white") +
     scale_fill_gradient(low="white", high="steelblue",
                         guide=guide_colourbar(title="IBS Score")) +
     labs(list(title="IBS Heat Map", x="Sample", y="Sample"))
p
```

